---
title: "Illegal Dumping in Philadelphia"
output: 
  flexdashboard::flex_dashboard:
    theme:
      bg: "#ffffff"
      fg: "#5E2C25"
      primary: "#C45B4D"
      base_font:
        google: Prompt
      code_font:
        google: JetBrains Mono
    orientation: columns
    vertical_layout: fill
---

```{=html}
<style type="text/css">
.value-box .value {
    color: white;
}
.value-box .caption {
    color: white;
}
</style>
```
```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(sf)
library(rphl)
library(plotly)
library(ggthemr)
library(tmap)
library(downloadthis)
library(RColorBrewer)
library(scales)
library(tigris)
library(monochromeR)
library(verbaliseR)


Sys.setenv(RSTUDIO_PANDOC="C:/Program Files/RStudio/resources/app/bin/quarto/bin/tools")

options(tigris_use_cache = TRUE)

tmap_options(basemaps = "Esri.WorldGrayCanvas") #set global tmap basemap

ggthemr("pale") #set global ggplot theme

choro_pal = rev(generate_palette("#C45B4D", modification = "go_lighter", n_colours = 5, view_palette = FALSE)) # create palette for choropleth map

blk_choro_pal = rev(generate_palette("#444444", modification = "go_lighter", n_colours = 5, view_palette = FALSE)) # create palette for choropleth map

map_pal = c("#3262AB", "#C45B4D","#444444")

status_pal = c("Open" = "#C45B4D",
                "Closed" = "#444444")

#swatch: "#444444" "#de6757" "#EB9050" "#3262AB" "#FF8D7D" "#C8E370" "#C45B4D" "#41a65c" "#5E2C25" "#78695F"
```

```{r data imports, cache = TRUE, include = FALSE}
base_url = "https://phl.carto.com/api/v2/sql"

one_year_ago = (lubridate::ymd(Sys.Date()) - lubridate::years(1))

query = sprintf("
        select *
        from public_cases_fc
        where requested_datetime  >= '%s'
        ", one_year_ago)
  
complaints = st_as_sf(get_carto(query,
                              format = 'csv',
                              base_url = base_url,
                              stringsAsFactors = FALSE) |>
                      filter(service_name == "Illegal Dumping",
                             !is.na(lat),
                             !is.na(lon)),
                      coords = c("lon", "lat"),
                      crs = st_crs('EPSG:4326')) |>
                      mutate(requested_datetime = as.Date(requested_datetime),
                             closed_datetime = as.Date(closed_datetime)) |>
                      st_transform(crs = st_crs("EPSG:2272")) # will need these to be projected for KDE later


complaints$closed_datetime[complaints$closed_datetime == ""] = NA

complaints = complaints |>
                              mutate(response_time_days = case_when(
                                                                is.na(closed_datetime) ~ as.numeric(difftime(Sys.Date(), requested_datetime, units = "days")),
                                                                TRUE ~ as.numeric(difftime(closed_datetime, requested_datetime, units = "days"))
                                                                  )
                                                                         )




phl_blocks = st_transform(tracts("PA", "Philadelphia"), crs = st_crs("EPSG:2272"))

phl_response_time = complaints |>
                        select(response_time_days) |>
                        aggregate(by = phl_blocks, FUN = mean)

all_comps_x_tract = complaints |>
                        aggregate(by = phl_blocks, FUN = length) |>
                        select(objectid) |>
                        rename(count = objectid)

open_comps_x_tract = complaints |>
                        filter(status == "Open") |>
                        aggregate(by = phl_blocks, FUN = length) |>
                        select(objectid) |>
                        rename(count = objectid)
```

## Column {data-width="300"}

### About the Dashboard

###### **What am I looking at?**

This dashboard displays illegal dumping complaints in Philadelphia since `r prettify_date((lubridate::ymd(Sys.Date()) - lubridate::years(1)), uk_or_us = "US")`. Click on the maps and graph to interact with them.

###### **Download Data**

To download a list of illegal dumping complaints in Philadelphia in the last year, click the button below:\
\

```{r download}
download_this((complaints |>
               select(address,
                      status,
                      subject,
                      service_name,
                      agency_responsible,
                      requested_datetime,
                      closed_datetime,
                      response_time_days
                      )),
    output_name = "Illegal Dumping Cases in Philadelphia",
    output_extension = ".xlsx",
    button_label = "Download Illegal Dumping Cases in Philadelphia",
    button_type = "warning",
    has_icon = TRUE,
    icon = "fa fa-save")
```

###### **Credits**

Dashboard built by [Nissim Lebovits](https://twitter.com/NissimLebovits). For questions, feedback, or requests, contact me at [nissim\@design.upenn.edu](mailto:nissim@design.upenn.edu){.email}.

### Current Open Cases

```{r current open tally}
valueBox(format(nrow(complaints |> filter(status == "Open")), big.mark = ","), icon = "fa-trash", color = "#C45B4D")
```

### Illegal Dumping Reports by Week

```{r density}
ggplotly(
  ggplot(complaints, aes(x = requested_datetime)) +
  geom_histogram(alpha = 0.3, fill = "#C45B4D", color = NA, bins = 52)+
    labs(y = "Count") +
  theme(legend.position = 'bottom',
        axis.title.x = element_blank())
    )
```

## Column {.tabset data-width="700"}

### Open Cases

*Current open illegal dumping cases*

```{r open map}
tmap_mode('view')

tm_shape(complaints |> filter(status == "Open")) +
  tm_dots(title = "Complaints",
          col = 'status',
          palette = status_pal,
          alpha = 0.2,
          id = "address",
              popup.vars = c("Status" = "status",
                             "Date Requested" = "requested_datetime",
                             "Date Closed" = "closed_datetime")) +
tm_view(view.legend.position = c("right", "bottom"),
        set.view = 12)
```

### Open Cases by Census Tract

*The number of current open cases per census tract*

```{r open cases}
tmap_mode('view')

tm_shape(open_comps_x_tract) +
  tm_polygons(
          title = "Open Cases",
          col = 'count',
          palette = choro_pal,
          style = "jenks",
          border.col = NA,
          border.alpha = 0,
          alpha = 0.7,
          id = "address",
          textNA = 'NA',
          popup.vars = c("Open Cases" = 'count')) +
tm_view(view.legend.position = c("right", "bottom"),
            colorNA = 'transparent',
        set.view = 12)
```

### Total Cases by Census Tract

*The number of total cases per census tract since `r prettify_date((lubridate::ymd(Sys.Date()) - lubridate::years(1)), uk_or_us = "US")`*

```{r total cases}
tmap_mode('view')

tm_shape(all_comps_x_tract) +
  tm_polygons(
          title = "Total Cases",
          col = 'count',
          palette = choro_pal,
          style = "jenks",
          border.col = NA,
          border.alpha = 0,
          alpha = 0.7,
          id = "address",
          textNA = 'NA',
          popup.vars = c("Total Cases" = 'count')) +
tm_view(view.legend.position = c("right", "bottom"),
            colorNA = 'transparent',
        set.view = 12)
```

### Average Response Time by Census Tract

*The average time (in days) for a complaint to be marked "closed" in each census tract*

```{r response time}
tmap_mode('view')

tm_shape(phl_response_time) +
  tm_polygons(
          title = "Avg. Response Time (Days)",
          col = 'response_time_days',
          palette = choro_pal,
          style = "jenks",
          border.col = NA,
          border.alpha = 0,
          alpha = 0.7,
          id = "address",
          textNA = 'NA',
          popup.vars = c("Avg. Response Time (Days)" = 'response_time_days')) +
tm_view(view.legend.position = c("right", "bottom"),
            colorNA = 'transparent',
        set.view = 12)
```
